---
title: "IMC Diabetes Damond"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, fig.path = "../Output/Figures/", dev=c('png','pdf','svg'))
```

## Load libraries

```{r message=FALSE}
library(tidyverse)
library(spicyR)
library(ggrepel)
```


# Load data

The segmented cells experiment can be created with "createSegmentedCellsDiabetes.R" file which includes instructions for downloading the data.

```{r}
load("../Data/cellExpDiabetes.RData")

ph <- imagePheno(cellExp)
ph$case <- factor(ph$case)

imagePheno(cellExp) <- ph

cellProp <- table(imageID(cellExp),cellType(cellExp))
dropCellTypes <- names(which(colMeans(cellProp==0)>0.5))


```


# Run spicyR using mixed effects model

```{r}
t1 <- Sys.time()


spicyTest <- spicy(cellExp, 
                   condition = "stage", 
                   subject = "case",
                   BPPARAM = BiocParallel::MulticoreParam(50),
                   includeZeroCells = FALSE,
                   Rs = seq(10,100,10),
                   weights = TRUE
                  )

t2 <- Sys.time()

t2 - t1

```


```{r}
top <- topPairs(spicyTest, n = 100, coef = 2)
top[top$from %in% c("alpha", "beta", "delta", "Th", "Tc", "neutrophil", "macrophage") & top$to %in% c("alpha", "beta", "delta", "Th", "Tc", "neutrophil", "macrophage"),]
```

```{r signifPlot, fig.height=5, fig.width=5}
signifPlot(spicyTest, breaks=c(-3, 3, 0.5), fdr = FALSE, marksToPlot = c("alpha", "beta", "delta", "Th", "Tc", "neutrophil", "macrophage"))

```

## View weights

```{r weights, fig.height=5, fig.width=7}

w <- spicyTest$weights
nCells <- spicyTest$nCells
m1 <- unlist(lapply(strsplit(names(w), "__"), function(x)x[1]))
m2 <- unlist(lapply(strsplit(names(w), "__"), function(x)x[2]))
counts1 <- as.vector(nCells[,m1])
counts2 <- as.vector(nCells[,m2])
df <- data.frame(counts1, counts2, weights = unlist(w))
ggplot(df, aes(sqrt(counts1),sqrt(counts2), colour = log10(weights)^(1/1))) + geom_point() + 
  theme_classic() + scale_color_continuous(type = "viridis") +
  xlim(0,65) + ylim(0,65)+
  labs(x = "sqrt ( cell type A )", y = "sqrt ( cell type B )", colour = "log10 ( weights )")

```

## Explore cytotoxix t cells vs beta cells relationship with diabetes.

```{r TcVsBeta, fig.height=5, fig.width=7}
df <- data.frame(u = spicyTest$pairwiseAssoc$Tc__beta, subject = imagePheno(cellExp)$case, stage = imagePheno(cellExp)$stage)
df <- dplyr::filter(df, stage %in% c("Onset", "Non-diabetic"))
ggplot(df, aes(x = paste(stage,"\n", "(", subject, ")", sep = ""), y = u)) + geom_hline(yintercept = 0, linetype = 2) + geom_boxplot() + theme_classic() + labs(x = "Stage\n(Case)", y = "u", title = "Tc vs beta cells") 

```



```{r TcVsBetaStage, fig.height=5, fig.width=2.5}
df <- data.frame(u = spicyTest$pairwiseAssoc$Tc__beta, subject = imagePheno(cellExp)$case, stage = imagePheno(cellExp)$stage)
df <- dplyr::filter(df, stage %in% c("Onset", "Non-diabetic"))
ggplot(df, aes(x = stage, y = u)) + geom_hline(yintercept = 0, linetype = 2) + geom_boxplot() + theme_classic() + labs(x = "Stage", y = "u", title = "Tc vs beta cells") 

```

## Explore Th cells relationship with diabetes.

```{r ThVsTh, fig.height=5, fig.width=7}
df <- data.frame(u = spicyTest$pairwiseAssoc$Th__Th, subject = imagePheno(cellExp)$case, stage = imagePheno(cellExp)$stage)
df <- filter(df, stage %in% c("Onset", "Non-diabetic"))
ggplot(df, aes(x = paste(stage,"\n", "(", subject, ")", sep = ""), y = u)) + geom_hline(yintercept = 0, linetype = 2) + geom_boxplot() + theme_classic() + labs(x = "Stage\n(Case)", y = "Localisation statistic: u", title = "Th vs Th cells") 

```

# Run spicyR using linear model

```{r}
t1 <- Sys.time()

spicyTestLM <- spicy(cellExp, 
                   condition = "stage", 
                   BPPARAM = BiocParallel::MulticoreParam(50),
                   includeZeroCells = FALSE,
                   Rs = seq(10,100,10),
                   weights = TRUE
                  )

t2 <- Sys.time()

t2 - t1

```


```{r}
top <- topPairs(spicyTestLM, n = 1000, coef = 2)
top[top$from %in% c("alpha", "beta", "delta", "Th", "Tc", "neutrophil", "macrophage") & top$to %in% c("alpha", "beta", "delta", "Th", "Tc", "neutrophil", "macrophage"),]
```


```{r signifPlotLM, fig.height=5, fig.width=5}
signifPlot(spicyTestLM, fdr = FALSE, marksToPlot = c("alpha", "beta","delta", "Th", "Tc", "neutrophil", "macrophage"), breaks = c(-10, 10, 0.5))

```

# Compare results

```{r lmVslmm, fig.height=6, fig.width=8}

df <- data.frame(LM = spicyTestLM$p.value[,2], LMM = spicyTest$p.value[,2], pairs = rownames(spicyTestLM$p.value))
df <- filter(df, spicyTestLM$comparisons$from%in%c("alpha", "beta", "delta", "Th", "Tc", "neutrophil", "macrophage") & spicyTestLM$comparisons$to%in%c("alpha", "beta", "delta", "Th", "Tc", "neutrophil", "macrophage"))


ggplot(df, aes(-log10(LMM), -log10(LM), label = pairs)) + geom_point() + theme_classic() + geom_abline(linetype = 2)

dfsub <- dplyr::filter(df, -log10(LMM) > 1.5 | -log10(LM) > 5)

ggplot(df, aes(-log10(LMM), -log10(LM), label = pairs)) + geom_point() + theme_classic() + geom_abline(linetype = 2) + geom_label_repel(data = dfsub[order(dfsub$LM),], size = 3, max.overlaps = 30, box.padding = 0.25) + ylim(-1, 18)

```


```{r}
sessionInfo()
```









